### 2022-12-14 hawaii/beakerBanter_hawaii.Rmd
Add Hawaii Dataset
Run Hawaii Dataset. Gitignore should include the binaries/database folders in the Hawaii folder dataset

---
title: "beakerbanter_hawaii"
author: "Shannon Rankin"
date: "2022-12-07"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```

## Data Processing for Beaked Whales: Hawaii

Start by loading the required packages

```{r, results=FALSE}
library(PAMpal)
library(banter)
library(rfPermute)
library(here)
here()
```

### PAMpal Processing

1.  Set up our PPS (PAMPal Settings Object) for the Hawaii dataset

```{r, eval=freshRun}
pps <- PAMpalSettings(db='Database/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  Process data and save to file to eliminate repeated processing.

***If*** this is the initial processing, ensure you have set 'freshRun =
TRUE' in code chunk at the top of this document so that you will run the
processPGDetections code and save the data as an rds file. In this case,
it will take some time to run, and we will check for warning messages
during processing.

```{r, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='hawaii_bw')
saveRDS(data, 'hawaii_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

***If*** you have already run the processing code, ensure you have set
'freshRun = FALSE' in code chunk at the top of this document so that you
will read in the existing .rds file and check for warning messages.

```{r, include=FALSE, eval=!freshRun}
data <- readRDS('hawaii_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

3.  Assign species identity according to original PAMguard labels, then
    relabel for consistency across projects.

```{r}
data <- setSpecies(data, 'pamguard')
reSpecies <- readRDS('species.rds')
print(reSpecies)
data <- setSpecies(data, 'reassign', value=reSpecies)
```

4.  Filter out unwanted data: (1) subset and rename species for BANTER
    model and (2) filter to retain only Channel 1.

```{r}
goodSpecies<- c("ZC", "MD", "BW", "BWC", "IP", "BW", "possBW")
data <- filter(data, species %in% goodSpecies)
data <- setSpecies(data, method='reassign',
                   value=data.frame(old=c('ZC', 'MD', 'BWC', 'IP'), new=c("Cuviers", "Blainsvilles", "CrossSeamount", "Indopacetus")))
data_ch1only <- filter(data, Channel == '1')
```

5.  Calculate Inter-Click Interval (ICI).

```{r}
data_ch1only <- calculateICI(data_ch1only, time='peakTime')
```

6.  Export data for BANTER (and drop species codes that will not be used
    for training). We will create two datasets: one with ICI and one
    without ICI, and save these for import into BANTER.

```{r, eval=freshRun}
banter_data <- export_banter(data_ch1only, dropSpecies = c('BW', 'possBW'), 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='hawaii_banter_data.rds')

banter_data_ici <- export_banter(data_ch1only, dropSpecies = c("BW", "possBW"), training=TRUE)
saveRDS(banter_data_ici, file='hawaii_banter_data_ici.rds')
```

### Build a BANTER Classification Model

#### *Model without ICI*

Load data, initialize a BANTER model, and create an initial BANTER
detector model (stage 1).

```{r, include=FALSE, eval=!freshRun}
banter_model <- readRDS('hawaii_banter_model_t5e3s10_t10e3s4.rds')
```

```{r, eval=freshRun}
banter_model <- initBanterModel(banter_data$events)
banter_model <- addBanterDetector(banter_model, banter_data$detectors, ntree=5e3, sampsize=10, importance = TRUE)
```

Check for stability (are more trees needed?) before moving on

```{r}
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 0:2))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 3:5))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 6))
summary(banter_model)
```

Run BANTER Event Model (stage 2)

```{r, eval=freshRun}
banter_model <- runBanterModel(banter_model, ntree=10e3, sampsize=4)
summary(banter_model)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r, eval=freshRun}
saveRDS(banter_model, 'hawaii_banter_model_t5e3s10_t10e3s4.rds')
```

#### *ICI Model*

Load data, initialize a BANTER model, and create an initial BANTER
detector model (stage 1).

```{r, include=FALSE, eval=!freshRun}
banter_model_ici <- readRDS('hawaii_banter_model_ici_t5e3s5_t10e3s4.rds')
```

```{r, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
# was 30e3 / 1 in SRs initial, too much memory. 5e3/1 is 210mb and takes ~10min
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=5e3, sampsize=5, importance = TRUE)
```

Check for stability to see if we need more trees

```{r}
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 0:2))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 3:5))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 6))
summary(banter_model_ici)
```

Run BANTER Event Model (stage 2)

```{r, eval=freshRun}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=10e3, sampsize=4)
summary(banter_model_ici)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r, eval=freshRun}
saveRDS(banter_model_ici, 'hawaii_banter_model_ici_t5e3s5_t10e3s4.rds')
```

### BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information)

First, identify the model you would like to examine (comment out the
model you [do not]{.underline} want to examine).

```{r}
#model <- banter_model_ici
model <- banter_model
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r}
banter_model_RF <- getBanterModel(model)
```

Confusion Matrix

```{r, include=TRUE}
hawaii_confuseMatrix <- confusionMatrix(banter_model_RF)
```

Class Priors (Expected Error Rate)

```{r, include=TRUE}
hawaii_priors <- classPriors(banter_model_RF, NULL)
```

Plot Confusion Matrix Heat Map

```{r, include=TRUE}
plotConfMat(banter_model_RF, title="ConfusionMatrix", plot=TRUE)
```

Plot Error Trace for Banter Model

```{r, include=TRUE}
plot(banter_model_RF)
```

PlotVotes

```{r, include=TRUE}
png("hawaii_votes.png", width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_RF)
dev.off()
```

Importance Heatmap

```{r, include=TRUE}
png("hawaii_importance.png", width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_RF, plot.type="heatmap")
dev.off()
```

Proximity Plot

```{r, include=TRUE}
png("hawaii_proximity.png", width = 20, height = 20, units = 'cm',  res = 300)
hawaii_proximityPlot <- plotProximity(banter_model_RF)
dev.off()
```

Plot Predicted Probabilities

```{r, include=TRUE}
plotPredictedProbs(banter_model_RF, bins=30, plot=TRUE)
```