### 2022-12-14 test.qmd

---
title: Open-Source machine learning BANTER acoustic classification of a cryptic echolocating species
author:
  - name: Shannon Rankin
    email: shannon.rankin@noaa.gov
    affiliations: 
        - id: SWFSC
          name: Southwest Fisheries Science Center
          department: Marine Mammal & Turtle Division
          address: 8901 La Jolla Shores Dr.
          city: La Jolla
          state: CA
          postal-code: 92037
    attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
  Passive acoustic monitoring drives me batty.
keywords: 
  - bioacoustics
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).