### 2023-02-07 hawaii/beakerBanter_hawaii.Rmd
Update manuscript and hawaii data
Extensive work on Manuscript and preparation of Hawaii data (in folder). Hawaii data *should* be good to knit. Manuscript has references, needs updated results, figures, inline data, and updated discussion.

---
title: "BANTER for Beaked Whales: Hawaii"
author: "Shannon Rankin"
date: "2022-02-06"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```

## PAMpal Data Processing

Start by loading the required packages

```{r, load packages, message=FALSE}
library(PAMpal)
library(banter)
library(rfPermute)
library(kableExtra)
library(here)
here()
```

1.  Set up our PPS (PAMPal Settings Object) for the Hawaii dataset

```{r, pampal settings, eval=freshRun}
pps <- PAMpalSettings(db='Database/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  Process data and save to file to eliminate repeated processing.

[If]{.underline} you have already run the processing code, ensure you
have set 'freshRun = FALSE' at top of this document to read in the
existing .rds file for downstream processing.

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data <- readRDS('hawaii_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

[If]{.underline} this is the [initial processing]{.underline}, ensure
you have set 'freshRun = TRUE' at top of this document to process and
save data. This will take some time to run.

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='hawaii_bw')
saveRDS(data, 'hawaii_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

3.  Assign species identity according to original PAMguard labels, then
    relabel for consistency across projects.

```{r, assign species ID}
data <- setSpecies(data, 'pamguard')
reSpecies <- readRDS('species.rds')
print(reSpecies)
data <- setSpecies(data, 'reassign', value=reSpecies)
```

3.  Filter out unwanted data: (1) subset and rename species for BANTER
    model and (2) filter to retain only Channel 1.

```{r, filter unwanted data}
goodSpecies<- c("ZC", "MD", "BW", "BWC", "IP", "BW", "possBW")
data <- filter(data, species %in% goodSpecies)
data <- setSpecies(data, method='reassign',
                   value=data.frame(old=c('ZC', 'MD', 'BWC', 'IP'), new=c("Cuviers", "Blainsvilles", "CrossSeamount", "Longmans")))
data_ch1only <- filter(data, Channel == '1')
```

5.  Calculate Inter-Click Interval (ICI).

```{r, calculate ICI}
data_ch1only <- calculateICI(data_ch1only, time='peakTime')
```

6.  Export data for BANTER (and drop species codes that will not be used
    for training). We will create two datasets: one with ICI and one
    without ICI, and save these for import into BANTER.

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data_ch1only, dropSpecies = c('BW', 'possBW'), 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='hawaii_banter_data.rds')

banter_data_ici <- export_banter(data_ch1only, dropSpecies = c("BW", "possBW"), training=TRUE)
saveRDS(banter_data_ici, file='hawaii_banter_data_ici.rds')
```

## Build a BANTER Classification Model

### *Model without ICI*

Load data, initialize a BANTER model, and create an initial BANTER
detector model (stage 1).

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model <- readRDS('hawaii_banter_model_t5e3s10_t10e3s4.rds')
```

```{r, banter detector model, eval=freshRun}
banter_model <- initBanterModel(banter_data$events)
banter_model <- addBanterDetector(banter_model, banter_data$detectors, ntree=5e3, sampsize=10, importance = TRUE)
```

Check for stability (are more trees needed?) before moving on

```{r, banter detector model stability}
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 0:2))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 3:5))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 6))
summary(banter_model)
```

Run BANTER Event Model (stage 2)

```{r, banter event model, eval=freshRun}
banter_model <- runBanterModel(banter_model, ntree=10e3, sampsize=4)
summary(banter_model)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r,save banter model, eval=freshRun}
saveRDS(banter_model, 'hawaii_banter_model_t5e3s10_t10e3s4.rds')
```

### *ICI Model*

Load data, initialize a BANTER model, and create an initial BANTER
detector model (stage 1).

```{r, read banter ici model, include=FALSE, eval=!freshRun}
banter_model_ici <- readRDS('hawaii_banter_model_ici_t5e3s5_t10e3s4.rds')
```

```{r, banter ici detector model, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=5e3, sampsize=5, importance = TRUE)
```

Check for stability to see if we need more trees

```{r, banter ici detector model stability}
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 0:2))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 3:5))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 6))
summary(banter_model_ici)
```

Run BANTER Event Model (stage 2)

```{r, banter ici event model, eval=freshRun}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=10e3, sampsize=4)
summary(banter_model_ici)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'hawaii_banter_model_ici_t5e3s5_t10e3s4.rds')
```

## BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information)

First, identify the model you would like to examine (comment out the
model you [do not]{.underline} want to examine).

```{r, identify model}
model <- banter_model
modelname <- "banter_model"

model_ici <- banter_model_ici
modelname_ici <- "banter_model_ici"
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r, extract RF model}
banter_model_RF <- getBanterModel(model)
banter_model_ici_RF <- getBanterModel(model_ici)
```

Confusion Matrix

```{r, confusion matrix, include=TRUE}
hawaii_confuseMatrix <- rfPermute::confusionMatrix(banter_model_RF)
hawaii_confuseMatrix <- kable(hawaii_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(5, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(5,hline_after = TRUE)%>%
  row_spec(5, bold = TRUE)%>%
  save_kable(paste(modelname, "hawaii_confuseMatrix.png", sep="_"), zoom = 2)

hawaii_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
hawaii_ici_confuseMatrix <- kable(hawaii_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(5, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(5,hline_after = TRUE)%>%
  row_spec(5, bold = TRUE)%>%
  save_kable(paste(modelname_ici, "hawaii_ici_confuseMatrix.png", sep="_"), zoom = 2)

```

Proximity Plot

```{r, proximity plot, include=TRUE}
png(paste(modelname, "hawaii_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
hawaii_proximityPlot <- plotProximity(banter_model_RF)
dev.off()
hawaii_proximityPlot <- plotProximity(banter_model_RF)

png(paste(modelname_ici, "hawaii_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
ici_hawaii_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_hawaii_proximityPlot <- plotProximity(banter_model_ici_RF)
```

Importance Heatmap

```{r, importance heat map, include=TRUE}
png(paste(modelname, "hawaii_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_RF, plot.type="heatmap")
dev.off()
hawaii_importance <- plotImportance(banter_model_RF, plot.type="heatmap")

png(paste(modelname_ici, "hawaii_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_hawaii_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")
```

PlotVotes

```{r, plot votes, include=TRUE}
png(paste(modelname, "hawaii_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_RF)
dev.off()
hawaii_votes <- plotVotes(banter_model_RF)

png(paste(modelname_ici, "hawaii_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_hawaii_votes <- plotVotes(banter_model_ici_RF)
```

Class Priors (Expected Error Rate)

```{r, class priors, include=TRUE}
hawaii_priors <- classPriors(banter_model_RF, NULL)
```

Plot Confusion Matrix Heat Map

```{r, confusion matrix heat map, include=TRUE}
plotConfMat(banter_model_RF, title="ConfusionMatrix", plot=TRUE)
```

Plot Error Trace for Banter Model

```{r, model error trace, include=TRUE}
plot(banter_model_RF)
```

Plot Predicted Probabilities

```{r, predicted probabilities, include=TRUE}
plotPredictedProbs(banter_model_RF, bins=30, plot=TRUE)
```