### 2023-02-07 natlantic/beakerBanter_natlantic.Rmd

---
title: 'BANTER for Beaked Whales: North Atlantic'
author: "Shannon Rankin"
date: "2023-01-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```

## PAMpal Data Processing

Start by loading the required packages

```{r, load packages, message=FALSE}
library (PAMpal)
library(banter)
library(rfPermute)
library(kableExtra)
library(readxl)
library(stringr)
library(here)
here()
```

1.  **Set up our PPS** (PAMPal Settings Object) for the NAtlantic
    dataset

```{r, pampal settings, eval=freshRun}
pps <- PAMpalSettings(db='Database/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  **Process data** and save to file to eliminate repeated processing.

[If]{.underline} you have already run the processing code, ensure you
have set 'freshRun = FALSE' at top of this document to read in the
existing .rds file for downstream processing.

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data <- readRDS('natlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

[If]{.underline} this is the [initial processing]{.underline}, ensure
you have set 'freshRun = TRUE' at top of this document to process and
save data. This will take some time to run.

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='natlantic_bw')
saveRDS(data, 'natlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

3.  **Assign species identity** according to original PAMguard labels,
    then relabel for consistency across projects.

    *Final species ID for beaked whale detections was provided by A.
    DeAngelis as an excel spreadsheet. The following code chunk pulls
    out the required information (event, species) in the format required
    for PAMpal.*

```{r, assign species ID}
HB1603_BW <- read_excel("HB1603_BWs.xlsx", sheet = "Alldata")#read in data
#create formatted event/species dataframe 
speciesID <- HB1603_BW[c("PGId", "Final Species Classification", "Database")] %>% 
  mutate("event" = paste(Database, PGId, sep=".OE")) %>% 
  rename("species" = "Final Species Classification") %>% 
  mutate("species" = str_replace_all(species, "'", ""),
         "species" = str_replace(species, "Gervais/Trues", "BW")) %>% 
  select(all_of(c("event", "species")))

#Assign species ID to data
data <- setSpecies(data, method='manual', value = speciesID)
#Note: 130 detections are not given species IDs and will be filtered out at a later step.
```

4.  **Filter out unwanted data**: Subset and rename species for BANTER
    model

```{r, filter unwanted data}
goodSpecies<- c(c("Trues", "Cuviers", "Sowerbys", "Gervais"))
data <- filter(data, species %in% goodSpecies)
```

5.  **Calculate Inter-Click Interval** (ICI).

```{r, calculate ICI}
data <- calculateICI(data, time='peakTime')
```

6.  **Export data for BANTER** (and drop species codes that will not be
    used for training). We will create two datasets: one with ICI and
    one without ICI, and save these for import into BANTER.

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data, dropSpecies = 'unid', 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='natlantic_banter_data.rds')

banter_data_ici <- export_banter(data, dropSpecies = 'unid', training=TRUE)
saveRDS(banter_data_ici, file='natlantic_banter_data_ici.rds')
```

## Build a BANTER Classification Model

### *Model without ICI*

1.  Load data, initialize a BANTER model, and create an initial BANTER
    detector model (stage 1).

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model <- readRDS('natlantic_banter_model_t5e3s3_t5e3s4.rds')
```

```{r, banter detector model, eval=freshRun}
banter_model <- initBanterModel(banter_data$events)
banter_model <- addBanterDetector(banter_model, banter_data$detectors, ntree=5e3, sampsize=3, importance = TRUE)
```

Check for stability (are more trees needed?) before moving on

```{r, banter detector model stability}
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 3:4))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 5:6))
summary(banter_model)
```

2.  Run BANTER Event Model (stage 2)

```{r, banter event model}
banter_model <- runBanterModel(banter_model, ntree=5e3, sampsize=4)
summary(banter_model)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r,save banter model, eval=freshRun}
saveRDS(banter_model, 'natlantic_banter_model_t5e3s3_t5e3s4.rds')
```

### *ICI Model*

1.  Load data, initialize a BANTER model, and create an initial BANTER
    detector model (stage 1).

```{r, read banter ici model, include=FALSE, eval=!freshRun}
banter_model_ici <- readRDS('natlantic_banter_model_ici_t10e3s2_t10e3s4.rds')
```

```{r, banter ici detector model, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=10e3, sampsize=2, importance = TRUE)
```

Check for stability to see if we need more trees

```{r, banter ici detector model stability}
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 3:4))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 5:6))
summary(banter_model_ici)
```

2.  Run BANTER Event Model (stage 2)

```{r, banter ici event model}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=10e3, sampsize=4)
summary(banter_model_ici)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'natlantic_banter_model_ici_t10e3s2_t10e3s4.rds')
```

## BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information)

First, identify the model you would like to examine (comment out the
model you [do not]{.underline} want to examine).

```{r, identify model}
model <- banter_model
modelname <- "banter_model"

model_ici <- banter_model_ici
modelname_ici <- "banter_model_ici"
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r, extract RF model, include=TRUE}
banter_model_RF <- getBanterModel(model)
banter_model_ici_RF <- getBanterModel(model_ici)
```

Confusion Matrix (simple model)

```{r, confusion matrix, include=TRUE}
natlantic_confuseMatrix <- rfPermute::confusionMatrix(banter_model_RF)
natlantic_confuseMatrix <- kable(natlantic_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(4, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(4, bold = TRUE)%>%
  save_kable(paste(modelname, "natlantic_confuseMatrix.png", sep="_"), zoom = 2)

natlantic_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
natlantic_ici_confuseMatrix <- kable(natlantic_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(4, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(4, bold = TRUE)%>%
  save_kable(paste(modelname_ici, "natlantic_ici_confuseMatrix.png", sep="_"), zoom = 2)

```

![](banter_model_natlantic_confuseMatrix.png)

Proximity Plot

```{r, proximity plot, include=TRUE}
png(paste(modelname, "natlantic_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
natlantic_proximityPlot <- plotProximity(banter_model_RF)
dev.off()
natlantic_proximityPlot <- plotProximity(banter_model_RF)

png(paste(modelname_ici, "natlantic_proximity.png", sep="_"), width = 20, height = 20, units = 'cm', res = 300)
ici_natlantic_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_natlantic_proximityPlot <- plotProximity(banter_model_ici_RF)
```

Importance Heatmap

```{r, importance heat map, include=TRUE}
png(paste(modelname, "natlantic_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_RF, plot.type="heatmap")
dev.off()
natlantic_importance <- plotImportance(banter_model_RF, plot.type="heatmap")

png(paste(modelname_ici, "natlantic_importance.png", sep="_"), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_natlantic_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")
```

PlotVotes

```{r, plot votes, include=TRUE}
png(paste(modelname, "natlantic_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_RF)
dev.off()
natlantic_votes <- plotVotes(banter_model_RF)

png(paste(modelname_ici, "natlantic_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_natlantic_votes <- plotVotes(banter_model_ici_RF)
```

Class Priors (Expected Error Rate)

```{r, class priors, include=TRUE}
natlantic_priors <- classPriors(banter_model_RF, NULL)
```

Plot Confusion Matrix Heat Map

```{r, confusion matrix heat map, include=TRUE}
plotConfMat(banter_model_RF, title="ConfusionMatrix", plot=TRUE)
```

Plot Error Trace for Banter Model

```{r, model error trace, include=TRUE}
plot(banter_model_RF)
```

Plot Predicted Probabilities

```{r, predicted probabilities, include=TRUE}
plotPredictedProbs(banter_model_RF, bins=30, plot=TRUE)
```