### 2023-03-02 epacific/beakerBanter_epacific.Rmd
add epacific analysis
Added updated epacific analysis using CCES

---
title: "BANTER for Beaked Whales: EPacific (CCES survey)"
author: "Shannon Rankin"
date: "2023-03-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```

## PAMpal Data Processing

Start by loading the required packages

```{r, load packages, message=FALSE}
library(PAMpal)
library(banter)
library(rfPermute)
library(kableExtra)
library(here)
here()
```

1.  **Set up our PPS** (PAMPal Settings Object) for the EPacific dataset

```{r, pampal settings, eval=freshRun}
pps <- PAMpalSettings(db='Databases/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  **Process data** and save to file to eliminate repeated processing.

[If]{.underline} you have already run the processing code, ensure you
have set 'freshRun = FALSE' at top of this document to read in the
existing .rds file for downstream processing.

```{r, read acoustic study, eval=!freshRun}
data <- readRDS('epacific_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

[If]{.underline} this is the [initial processing]{.underline}, ensure
you have set 'freshRun = TRUE' at top of this document to process and
save data. This will take some time to run.

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='epacific_bw')
saveRDS(data, 'epacific_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

3.  **Add GPS data**: Add GPS data from PAMGuard table

    ``` {r, add GPS}
    addGps(data)
    getWarnings(data)
    ```

4.  **Assign species identity** according to original PAMguard labels,
    then relabel for consistency across projects.

```{r, assign species ID}
data <- setSpecies(data, 'pamguard')
reSpecies <- readRDS('species.rds')
print(reSpecies)
data <- setSpecies(data, 'reassign', value=reSpecies)
```

5.  **Filter out unwanted data**: (1) subset and rename species for
    BANTER model and (2) filter to retain only Channel 1.

```{r, filter unwanted data}
goodSpecies<- c("ZC", "MS", "BB", "BW", "BW37V", "BW43", "possBW", "BWC")
data <- filter(data, species %in% goodSpecies)
data <- setSpecies(data, method='reassign',
                   value=data.frame(old=c('ZC', 'MS', 'BB', 'BWC'), new=c("Cuviers", "Stejnegers", "Bairds", "CrossSeamount")))
data_ch1only <- filter(data, Channel == '1')
```

6. **Calculate Inter-Click Interval** (ICI).

```{r, calculate ICI}
data_ch1only <- calculateICI(data_ch1only, time='peakTime')
```

7. **Export Event Wav Files** Wav files for events will be exported to a local drive, and then uploaded to Figshare.

``` {r, export events wav, eval=freshRun}
addRecordings (data_ch1only, log = TRUE )

wavDir <- "./ccesEventWav" #identify directory to write clips to
writeEventClips(data_ch1only, buffer = 5, mode = 'event', useSample = TRUE, folder = wavDir)
#Once this is finished, this data needs to be uploaded to a project in Figshare. Obtain the article ID and post below in annomate section

```

8. **Export to Annomate** Export annotation information to a csv for upload to Ann-O-Mate.
**NOTE** DO NOT SAVE THIS CODE CHUNK PUBLICLY--

```{r, export to annomate, include=TRUE, eval=freshRun}
#First, run:
unique(species(data)) #to see unique species and order in which they are named
#Then update line 119 to include full species names in order in which they occur
specMap <- data.frame(
    old = unique(species(data)),
    new = NA)
specMap$new <- c('Possible Beaked Whale', 'Ziphius cavirostris', 'Mesoplodon stejnegeri', 
                 'BW39V', 'Unidentified Beaked Whale', 'Berardius bairdii', 'BW43')
anno <- prepAnnotation(data, specMap = specMap, mode='event',
                       # following args are optional, fill in
                       # any that will be the same for all entries
                       source = 'figshare.com',
                       contact = 'shannon.rankin@noaa.gov')
#Export Annomate file, does this info save to acoustic study?

# to get figshare data you will need the article id and your personal token
# i recommend storing your token in an external file instead of typing it in here
# also DO NOT commit this file to any repository
figToken <- yaml::read_yaml('secrets.yaml')$fig_token
figId <- 13393360
figshareData <- getFigshareInfo(figToken, figId)

anno <- matchRecordingUrl(anno, figshareData)
# will give you warnings and messages about missing fields
checkAnnotation(anno)
# if you need to fix any manually, write to csv then read back in
annoFile <- 'AnnomateExport.csv'
write.csv(anno, file=annoFile, row.names = FALSE)
# go fix stuff
anno <- read.csv(annoFile, stringsAsFactors = FALSE)
# recheck to see if all seems fine
checkAnnotation(anno)
# add to acoustic study for storage
data <- addAnnotation(data, anno)
# this creates CSV ready for figshare, will also repeat messages from the check
export_annomate(data, file=annoFile)

```

9. Export data for BANTER (and drop species codes that will not be used
    for training). We will create two datasets: one with ICI and one
    without ICI, and save these for import into

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data_ch1only, dropSpecies = c('BW', 'possBW'), 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='epacific_banter_data.rds')

banter_data_ici <- export_banter(data_ch1only, dropSpecies = c("BW", "possBW"), training=TRUE)
saveRDS(banter_data_ici, file='epacific_banter_data_ici.rds')

#save update of Acoustic Study
saveRDS(data_ch1only, 'epacific_study.rds')
saveRDS(data, 'epacificALL_study.rds')
```

## Build a BANTER Classification Model

### *Simple Model*

Load data, initialize a BANTER model, and create an initial BANTER
detector model (stage 1).

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model <- readRDS('epacific_banter_model_t1e4s3_t1e4s4.rds')
```

```{r, banter detector model, eval=freshRun}
banter_model <- initBanterModel(banter_data$events)
banter_model <- addBanterDetector(banter_model, banter_data$detectors, ntree=1e4, sampsize=3, importance = TRUE)
```

Check for stability (are more trees needed?) before moving on

```{r, banter detector model stability, eval=freshRun}
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 1:3))
plotDetectorTrace(banter_model, detector = paste0('Click_Detector_', 4:6))
summary(banter_model)
```

Run BANTER Event Model (stage 2)

```{r, banter event model, eval=freshRun}
banter_model <- runBanterModel(banter_model, ntree=1e4, sampsize=4)
summary(banter_model)
```

Check for stability (are more trees needed?). If acceptable, save model
with tree/sampsize info in the filename.

```{r,save banter model, eval=freshRun}
saveRDS(banter_model, 'epacific_banter_model_t1e4s3_t1e4s4.rds')
```

### *ICI Model*

Load data (if previously run)

```{r, read banter ici model, include=FALSE, eval=!freshRun}
banter_model_ici <- readRDS('epacific_banter_model_ici_t1e4s3_t1e4s4.rds')
```

Fresh Run: Initialize, Run & Evaluate Detector Model

```{r, banter ici detector model, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=1e4, sampsize=3, importance = TRUE)
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 1:3))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 4:6))
summary(banter_model_ici)
```

Run Event Model

```{r, banter ici event model, eval=freshRun}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=1e4, sampsize=4)
summary(banter_model_ici)
```

Save BANTER model

```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'epacific_banter_model_ici_t1e4s3_t1e4s4.rds')
```

## BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information. First, identify the model you would like to examine
(comment out the model you [do not]{.underline} want to examine).

```{r, identify model}
model <- banter_model
modelname <- "banter_model"

model_ici <- banter_model_ici
modelname_ici <- "banter_model_ici"
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r, extract RF model}
banter_model_RF <- getBanterModel(model)
banter_model_ici_RF <- getBanterModel(model_ici)
```

Confusion Matrix

```{r, confusion matrix, include=TRUE}
epacific_confuseMatrix <- rfPermute::confusionMatrix(banter_model_RF)
epacific_confuseMatrix <- kable(epacific_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(6, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(4,hline_after = TRUE)%>%
  row_spec(6, bold = TRUE)%>%
  save_kable(paste(modelname, "epacific_confuseMatrix.png", sep="_"), zoom = 2)

epacific_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
epacific_ici_confuseMatrix <- kable(epacific_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(6, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(4,hline_after = TRUE)%>%
  row_spec(6, bold = TRUE)%>%
  save_kable(paste(modelname_ici, "epacific_ici_confuseMatrix.png", sep="_"), zoom = 2)
```

Proximity Plot

```{r, proximity plot, include=TRUE}
png(paste(modelname, "epacific_proximity.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
epacific_proximityPlot <- plotProximity(banter_model_RF)
dev.off()
epacific_proximityPlot <- plotProximity(banter_model_RF)

png(paste(modelname_ici, "epacific_proximity.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
ici_epacific_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_epacific_proximityPlot <- plotProximity(banter_model_ici_RF)

```

Importance Heatmap

```{r, importance heat map, include=TRUE}
png(paste(modelname, "epacific_importance.png", sep="_"), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_RF, plot.type="heatmap")
dev.off()
epacific_importance <- plotImportance(banter_model_RF, plot.type="heatmap")

png(paste(modelname_ici, "epacific_importance.png", sep="_"), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_epacific_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")
```

PlotVotes

```{r, plot votes, include=TRUE}
png(paste(modelname, "epacific_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_RF)
dev.off()
epacific_votes <- plotVotes(banter_model_RF)

png(paste(modelname_ici, "epacific_votes.png", sep="_"), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_epacific_votes <- plotVotes(banter_model_ici_RF)
```

Class Priors (Expected Error Rate)

```{r, class priors, include=TRUE}
epacific_priors <- classPriors(banter_model_RF, NULL)
```

Plot Confusion Matrix Heat Map

```{r, confusion matrix heat map, include=TRUE}
plotConfMat(banter_model_RF, title="ConfusionMatrix", plot=TRUE)
```

Plot Error Trace for Banter Model

```{r, model error trace, include=TRUE}
plot(banter_model_RF)
```

Plot Predicted Probabilities

```{r, predicted probabilities, include=TRUE}
plotPredictedProbs(banter_model_RF, bins=30, plot=TRUE)
```