### 2023-06-08 epacific/beakerBanter_epacific.Rmd
Major revisions for draft
Revised all datasets and created draft manuscript

---
title: "BANTER for Beaked Whales: EPacific (CCES survey)"
author: "Shannon Rankin"
date: "2023-05-25"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```

###Organizational Instructions Save this file in an R Project that
includes the following foldering: /Project \|-Database/ \|-Binaries/
\|-Figures/ \|-epacific \|\_ beakerBanter_epacific.Rmd \|\_
epacific_banter_data.rds \|\_ epacific_banter_data_ici.rds \|\_
epacific_banter_model_ec_t1e4s3_t1e4s4.rds \|\_
epacific_banter_model_ici_t1e4s3_t1e4s4

## PAMpal Data Processing

Start by loading the required packages

```{r, load packages, message=FALSE}
library("easypackages")
libraries("PAMpal", "banter", "rfPermute", "kableExtra", "magick", "magrittr", "here")
here()
```

1.  **Set up our PPS** (PAMPal Settings Object) for the EPacific dataset

```{r, pampal settings, eval=freshRun}
pps <- PAMpalSettings(db='Databases/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  **Process data** and save to file to eliminate repeated processing.

[If]{.underline} this is the [initial processing]{.underline}, ensure
you have set 'freshRun = TRUE' at top of this document to process and
save data. This will take some time to run.

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='epacific_bw')
saveRDS(data, 'epacific_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

[If]{.underline} you have already run the processing code, ensure you
have set 'freshRun = FALSE' at top of this document to read in the
existing .rds file for downstream processing.

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data_ch1only <- readRDS('epacific_study.rds')
# Double check warning messages
print(getWarnings(data_ch1only)$message)#Note: Detection >7200sec apart is a sperm whale detection. Anne was informed of the error.
```

3.  **Assign species identity** according to original PAMguard labels,
    then relabel for consistency across projects.

```{r, assign species ID, eval=freshRun}
data <- setSpecies(data, 'pamguard')
reSpecies <- readRDS('species.rds')
print(reSpecies)
data <- setSpecies(data, 'reassign', value=reSpecies)
```

4.  **Filter out unwanted data**: (1) subset and rename species for
    BANTER model and (2) filter to retain only Channel 1.

```{r, filter unwanted data, eval=freshRun}
data <- setSpecies(data, method='reassign',
                   value=data.frame(old=c('ZC', 'MS', 'BB', 'BWC', 'BW37V'), new=c("Cuviers", "Stejnegers", "Bairds", "CrossSeamount", "Hubbs")))
goodSpecies<- c("Cuviers", "Stejnegers", "Bairds", "CrossSeamount", "Hubbs", "BW", "BW43", "possBW")
data <- filter(data, species %in% goodSpecies)
data_ch1only <- filter(data, Channel == '1')
```

5.  **Calculate Inter-Click Interval** (ICI).

```{r, calculate ICI, eval=freshRun}
data_ch1only <- calculateICI(data_ch1only, time='peakTime')
```

6.  **Add GPS data**: Add GPS data from PAMGuard table (2 hour
    threshold), then filter out data without GPS

```{r, add GPS, eval=freshRun}
data_ch1only <- addGps(data_ch1only, thresh = 7200)#11 events over the threshold
getWarnings(data_ch1only)

data_ch1only <- filter(data_ch1only, !is.na(Latitude))#filter out events w/o GPS
clicks <- getClickData(data_ch1only)
sum(is.na(clicks$Latitude))
unique(clicks[is.na(clicks$Latitude), 'eventId'])#confirm there are no remaining events w/o GPS
```

7. **Add Environmental Data**
```{r, Add Environmental Data, eval=freshRun}
#SST
data_ch1only <- matchEnvData(data_ch1only, nc='jplMURSST41mday', var='sst')
#Seafloor Depth
data_ch1only <- matchEnvData(data_ch1only, nc='erdSrtm30plusSeafloorGradient', var='sea_floor_depth')
#Seafloor Gradient
data_ch1only <- matchEnvData(data_ch1only, nc='erdSrtm30plusSeafloorGradient', var='magnitude_gradient')

```

7.  **Export Event Wav Files** Wav files for events will be exported to
    a local drive, and then uploaded to Figshare.

```{r, export events wav, eval=freshRun}
#This should only be run one time, with access to original wav files. Will be commented out to ensure it is not accidentally run
#data <- readRDS(here('epacific', 'epacific_study.rds'))
#Check that all events are <120s
#checkStudy(data, maxLength = 120)

#Identify location of all recording files and log files
#data<- addRecordings (data, log = TRUE )
#saveRDS(data, 'epacific_study.rds')

#Identify directory to save wav files, and write event Wav Clips
#wavDir <- here('epacific', 'ccesEventWav')#identify directory to write clips to
#writeEventClips(data, buffer = 5, mode = 'event', useSample = FALSE, outDir = wavDir)
#saveRDS(data, 'epacific_study.rds')

```

8.  **Export to Annomate** Export annotation information to a csv for
    upload to Ann-O-Mate. Run code in annomate.R, but do not share this code publicly.

9.  Export data for BANTER (and drop species codes that will not be used
    for training). We will create two datasets: one with ICI and one
    without ICI, and save these for import into

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data_ch1only, dropSpecies = c('BW', 'possBW'), 
                         dropVars = c('All_ici', 'Click_Detector_3_ici', 'sea_floor_depth_mean', 'sst_mean', 'productivity_mean', 'chlorophyll_mean',  'magnitude_gradient_mean'), training=TRUE)
saveRDS(banter_data, file='epacific_banter_data.rds')

banter_data_ici <- export_banter(data_ch1only, dropSpecies = c("BW", "possBW"), dropVars = c('sea_floor_depth_mean', 'sst_mean', 'productivity_mean', 'chlorophyll_mean',  'magnitude_gradient_mean'),training=TRUE)
saveRDS(banter_data_ici, file='epacific_banter_data_ici.rds')

banter_data_env <- export_banter(data_ch1only, dropSpecies = c("BW", "possBW"), training=TRUE)
saveRDS(banter_data_env, file='epacific_banter_data_env.rds')

#save update of Acoustic Study
saveRDS(data_ch1only, 'epacific_study.rds')
saveRDS(data, 'epacificALL_study.rds')
```

## Build a BANTER Classification Model

### *EC (only) Model*

Initialize, Run & Evaluate Detector Model (stage 1).

```{r, banter ec detector model, eval=freshRun}
banter_model_ec <- initBanterModel(banter_data$events)
banter_model_ec <- addBanterDetector(banter_model_ec, banter_data$detectors, ntree=1e4, sampsize=3, importance = TRUE)

plotDetectorTrace(banter_model_ec, detector = paste0('Click_Detector_', 1:3))
plotDetectorTrace(banter_model_ec, detector = paste0('Click_Detector_', 4:6))
summary(banter_model_ec)
```

Run BANTER Event Model (stage 2)

```{r, banter ec event model, eval=freshRun}
banter_model_ec <- runBanterModel(banter_model_ec, ntree=1e4, sampsize=4)
summary(banter_model_ec)
```

Once a stable model is identified, save model with tree/sampsize info in
the filename.

```{r,save banter ec model, eval=freshRun}
saveRDS(banter_model_ec, 'epacific_banter_model_ec_t1e4s3_t1e4s4.rds')
```

### *ICI Model*

Initialize, Run & Evaluate Detector Model (stage 1).

```{r, banter ici detector model, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=1e4, sampsize=3, importance = TRUE)

plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 1:3))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 4:6))
summary(banter_model_ici)
```

Run Event Model (stage 2)

```{r, banter ici event model, eval=freshRun}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=1e4, sampsize=4)
summary(banter_model_ici)
```

Once a stable model is identified, save model with tree/sampsize info in
the filename.

```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'epacific_banter_model_ici_t1e4s3_t1e4s4.rds')
```
### *ENV Model*

Initialize, Run & Evaluate Detector Model (stage 1).
```{r, banter env detector model, eval=freshRun}
banter_model_env <- initBanterModel(banter_data_env$events)
banter_model_env <- addBanterDetector(banter_model_env, banter_data_env$detectors, ntree=1e4, sampsize=1, importance = TRUE)

plotDetectorTrace(banter_model_env, detector = paste0('Click_Detector_', 1:3))
plotDetectorTrace(banter_model_env, detector = paste0('Click_Detector_', 4:6))
summary(banter_model_env)
```

Run Event Model (stage 2)
```{r, banter env event model, eval=freshRun}
banter_model_env <- runBanterModel(banter_model_env, ntree=1e4, sampsize=5)
summary(banter_model_env)
```

Once a stable model is identified, save model with tree/sampsize info in
the filename.

```{r, save banter env model, eval=freshRun}
saveRDS(banter_model_env, 'epacific_banter_model_env_t1e4s1_t1e4s5.rds')
```
## BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information.

First, load the models (if not a fresh run)

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model_ec <- readRDS('epacific_banter_model_ec_t1e4s3_t1e4s4.rds')
banter_model_ici <- readRDS('epacific_banter_model_ici_t1e4s3_t1e4s4.rds')
banter_model_env <- readRDS('epacific_banter_model_env_t1e4s1_t1e4s5.rds')
```

Identify the model you would like to examine (comment out the model you
[do not]{.underline} want to examine).

```{r, identify model, eval=!freshRun}
model_ec <- banter_model_ec
modelname_ec <- "banter_model_ec"

model_ici <- banter_model_ici
modelname_ici <- "banter_model_ici"

model_env <- banter_model_env
modelname_env <- "banter_model_env"
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r, extract RF model, eval=!freshRun}
banter_model_ec_RF <- getBanterModel(model_ec)
banter_model_ici_RF <- getBanterModel(model_ici)
banter_model_env_RF <- getBanterModel(model_env)
```

Class Priors (Expected Error Rate)

```{r, class priors, include=TRUE, eval=!freshRun}
epacific_ec_priors <- classPriors(banter_model_ec_RF, NULL)[,1]
epacific_ici_priors <- classPriors(banter_model_ici_RF, NULL)[,1]
epacific_env_priors <- classPriors(banter_model_env_RF, NULL)[,1]
```

Confusion Matrix

```{r, confusion matrix, include=TRUE, eval=!freshRun}
epacific_ec_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ec_RF)
epacific_ec_confuseMatrix <- cbind(epacific_ec_confuseMatrix, priors = epacific_ec_priors)
epacific_ec_confuseMatrix <- kable(epacific_ec_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
   column_spec(7, border_right = TRUE)%>%
   row_spec(0, bold = TRUE)%>%
   row_spec(4,hline_after = TRUE)%>%
   row_spec(7, bold = TRUE)%>%
   save_kable("../manuscript/manuscript_files/epacific_ec_confuseMatrix.png", zoom = 9)

epacific_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
epacific_ici_confuseMatrix <- cbind(epacific_ici_confuseMatrix, priors = epacific_ici_priors)
epacific_ici_confuseMatrix <- kable(epacific_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(7, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(4,hline_after = TRUE)%>%
  row_spec(7, bold = TRUE)%>%
  save_kable("../manuscript/manuscript_files/epacific_ici_confuseMatrix.png", zoom = 9)

epacific_env_confuseMatrix <- rfPermute::confusionMatrix(banter_model_env_RF)
epacific_env_confuseMatrix <- cbind(epacific_env_confuseMatrix, priors = epacific_env_priors)
epacific_env_confuseMatrix <- kable(epacific_env_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(7, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(4,hline_after = TRUE)%>%
  row_spec(7, bold = TRUE)%>%
  save_kable("../manuscript/manuscript_files/epacific_env_confuseMatrix.png", zoom = 9)
```

![BANTER Model E.Pacific EC Confusion
Matrix](../manuscript/manuscript_files/epacific_ec_confuseMatrix.png)

![BANTER Model E.Pacific ICI Confusion
Matrix](../manuscript/manuscript_files/epacific_ici_confuseMatrix.png)

![BANTER Model E.Pacific ENV Confusion
Matrix](../manuscript/manuscript_files/epacific_env_confuseMatrix.png)

Proximity Plot

```{r, proximity plot, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/epacific_ec_proximity.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotProximity(banter_model_ec_RF)
dev.off()
ec_epacific_proximityPlot <- plotProximity(banter_model_ec_RF)

png(('../manuscript/manuscript_files/epacific_ici_proximity.png'), width = 20, height = 20, units = 'cm',  res = 300)
ici_epacific_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_epacific_proximityPlot <- plotProximity(banter_model_ici_RF)

png(('../manuscript/manuscript_files/epacific_env_proximity.png'), width = 20, height = 20, units = 'cm',  res = 300)
env_epacific_proximityPlot <- plotProximity(banter_model_env_RF)
dev.off()
env_epacific_proximityPlot <- plotProximity(banter_model_env_RF)
```

Importance Heatmap

```{r, importance heat map, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/epacific_ec_importance.png'), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_ec_RF, plot.type="heatmap")
dev.off()
ec_epacific_importance <- plotImportance(banter_model_ec_RF, plot.type="heatmap")

png(('../manuscript/manuscript_files/epacific_ici_importance.png'), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_epacific_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")

png(('../manuscript/manuscript_files/epacific_env_importance.png'), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_env_RF, plot.type="heatmap")
dev.off()
env_epacific_importance <- plotImportance(banter_model_env_RF, plot.type="heatmap")
```

PlotVotes

```{r, plot votes, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/epacific_ec_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ec_RF)
dev.off()
epacific_votes <- plotVotes(banter_model_ec_RF)

png(('../manuscript/manuscript_files/epacific_ici_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_epacific_votes <- plotVotes(banter_model_ici_RF)

png(('../manuscript/manuscript_files/epacific_env_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_env_RF)
dev.off()
env_epacific_votes <- plotVotes(banter_model_env_RF)
```

Plot Predicted Probabilities

```{r, predicted probabilities, include=TRUE}
plotPredictedProbs(banter_model_ec_RF, bins=30, plot=TRUE)
plotPredictedProbs(banter_model_ici_RF, bins=30, plot=TRUE)
plotPredictedProbs(banter_model_env_RF, bins=30, plot=TRUE)
```

Create Figure for Publication

```{r, Publication Figures, include=TRUE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

epacific_ec_Figure <-image_append(c(prox, heat, vote))
epacific_ec_Figure<- image_append(c(confuse, epacific_ec_Figure), stack=TRUE)
image_write(epacific_ec_Figure, path = here('manuscript', 'manuscript_files','epacific_ec_Figure.png'), format ='png')
print(epacific_ec_Figure, info=FALSE)

confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

epacific_ici_Figure <-image_append(c(prox, heat, vote))
epacific_ici_Figure<- image_append(c(confuse, epacific_ici_Figure), stack=TRUE)
image_write(epacific_ici_Figure, path = here('manuscript', 'manuscript_files','epacific_ici_Figure.png'), format ='png')
print(epacific_ici_Figure, info=FALSE)

confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_env_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_env_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_env_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_env_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

epacific_env_Figure <-image_append(c(prox, heat, vote))
epacific_env_Figure<- image_append(c(confuse, epacific_env_Figure), stack=TRUE)
image_write(epacific_env_Figure, path = here('manuscript', 'manuscript_files','epacific_env_Figure.png'), format ='png')
print(epacific_env_Figure, info=FALSE)
```