### 2023-06-08 hawaii/beakerBanter_hawaii.Rmd

---
title: "BANTER for Beaked Whales: Hawaii"
author: "Shannon Rankin"
date: "2023-05-24"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```
/Project
  |-Database/
  |-Binaries/
  |-Figures/
  |-hawaii
    |_ beakerBanter_hawaii.Rmd
    |_ hawaii_banter_data.rds
    |_ hawaii_banter_data_ici.rds
    |_ hawaii_banter_model_t5e3s10_t1e4s4.rds
    |_ hawaii_banter_model_ici_t5e3s5_t1e4s4.rds

## PAMpal Data Processing

Start by loading the required packages

```{r, load packages, message=FALSE}
library("easypackages")
libraries("PAMpal", "banter", "rfPermute", "kableExtra", "magick", "magrittr", "here")
here()
```

1.  Set up our PPS (PAMPal Settings Object) for the Hawaii dataset

```{r, pampal settings, eval=freshRun}
pps <- PAMpalSettings(db='Database/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  Process data and save to file to eliminate repeated processing.

[If]{.underline} this is the [initial processing]{.underline}, ensure
you have set 'freshRun = TRUE' at top of this document to process and
save data. This will take some time to run.

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='hawaii_bw')
saveRDS(data, 'hawaii_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

[If]{.underline} you have already run the processing code, ensure you
have set 'freshRun = FALSE' at top of this document to read in the
existing .rds file for downstream processing.

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data_ch1only <- readRDS('hawaii_study.rds')
# Double check warning messages
print(getWarnings(data_ch1only)$message)
```

3.  Assign species identity according to original PAMguard labels, then
    relabel for consistency across projects.

```{r, assign species ID, eval=freshRun}
data <- setSpecies(data, 'pamguard')
reSpecies <- readRDS('species.rds')
print(reSpecies)
data <- setSpecies(data, 'reassign', value=reSpecies)
```

4.  Filter out unwanted data: (1) subset and rename species for BANTER
    model and (2) filter to retain only Channel 1.

```{r, filter unwanted data, eval=freshRun}
goodSpecies<- c("ZC", "MD", "BW", "BWC", "IP", "BW", "possBW")
data <- filter(data, species %in% goodSpecies)
data <- setSpecies(data, method='reassign',
                   value=data.frame(old=c('ZC', 'MD', 'BWC', 'IP'), new=c("Cuviers", "Blainsvilles", "CrossSeamount", "Longmans")))
data_ch1only <- filter(data, Channel == '1')
```

5.  Calculate Inter-Click Interval (ICI).

```{r, calculate ICI, eval=freshRun}
data_ch1only <- calculateICI(data_ch1only, time='peakTime')
```

6.  Export data for BANTER (and drop species codes that will not be used
    for training). We will create two datasets: one with ICI and one
    without ICI, and save these for import into BANTER.

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data_ch1only, dropSpecies = c('BW', 'possBW'), 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='hawaii_banter_data.rds')

banter_data_ici <- export_banter(data_ch1only, dropSpecies = c("BW", "possBW"), training=TRUE)
saveRDS(banter_data_ici, file='hawaii_banter_data_ici.rds')

#save update of Acoustic Study
saveRDS(data_ch1only, 'hawaii_study.rds')
```

## Build a BANTER Classification Model

### *EC (only) Model*

Initialize, Run & Evaluate Detector Model (stage 1).
```{r, banter ec detector model, eval=freshRun}
banter_model_ec <- initBanterModel(banter_data$events)
banter_model_ec <- addBanterDetector(banter_model_ec, banter_data$detectors, ntree=5e3, sampsize=10, importance = TRUE)

plotDetectorTrace(banter_model_ec, detector = paste0('Click_Detector_', 0:3))
plotDetectorTrace(banter_model_ec, detector = paste0('Click_Detector_', 4:6))
summary(banter_model_ec)
```

Run BANTER Event Model (stage 2)
```{r, banter ec event model, eval=freshRun}
banter_model_ec <- runBanterModel(banter_model_ec, ntree=1e4, sampsize=4)
summary(banter_model_ec)
```

Once a stable model is identified, save model with tree/sampsize info in the filename.
```{r,save banter ec model, eval=freshRun}
saveRDS(banter_model_ec, 'hawaii_banter_model_ec_t5e3s10_t1e4s4.rds')
```

### *ICI Model*

Initialize, Run & Evaluate Detector Model (stage 1)
```{r, banter ici detector model, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=5e3, sampsize=5, importance = TRUE)

plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 0:3))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 4:6))
summary(banter_model_ici)
```

Run BANTER Event Model (stage 2)
```{r, banter ici event model, eval=freshRun}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=1e4, sampsize=4)
summary(banter_model_ici)
```

Once a stable model is identified, save model with tree/sampsize info in the filename.
```{r, save banter ici model, eval=freshRun}
saveRDS(banter_model_ici, 'hawaii_banter_model_ici_t5e3s5_t1e4s4.rds')
```

## BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information)

```{r, read banter model, include=FALSE, eval=!freshRun}
banter_model_ec <- readRDS('hawaii_banter_model_ec_t5e3s10_t1e4s4.rds')
banter_model_ici <- readRDS('hawaii_banter_model_ici_t5e3s5_t1e4s4.rds')
```

First, identify the model you would like to examine (comment out the
model you [do not]{.underline} want to examine).

```{r, identify model}
model_ec <- banter_model_ec
modelname_ec <- "banter_model_ec"

model_ici <- banter_model_ici
modelname_ici <- "banter_model_ici"
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r, extract RF model}
banter_model_ec_RF <- getBanterModel(model_ec)
banter_model_ici_RF <- getBanterModel(model_ici)
```

Class Priors (Expected Error Rate)

```{r, class priors, include=TRUE}
hawaii_ec_priors <- classPriors(banter_model_ec_RF, NULL)[,1]
hawaii_ici_priors <- classPriors(banter_model_ici_RF, NULL)[,1]
```

Confusion Matrix

```{r, confusion matrix, include=TRUE}
hawaii_ec_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ec_RF)
hawaii_ec_confuseMatrix <- cbind(hawaii_ec_confuseMatrix, priors = hawaii_ec_priors)
hawaii_ec_confuseMatrix <- kable(hawaii_ec_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(5, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(5,hline_after = TRUE)%>%
  row_spec(5, bold = TRUE)%>%
  save_kable('../manuscript/manuscript_files/hawaii_ec_confuseMatrix.png', zoom = 9)

hawaii_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
hawaii_ici_confuseMatrix <- cbind(hawaii_ici_confuseMatrix, priors = hawaii_ici_priors)
hawaii_ici_confuseMatrix <- kable(hawaii_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(5, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(5,hline_after = TRUE)%>%
  row_spec(5, bold = TRUE)%>%
  save_kable('../manuscript/manuscript_files/hawaii_ici_confuseMatrix.png', zoom = 9)

```

![BANTER Model Hawaii EC Confusion
Matrix](../manuscript/manuscript_files/hawaii_ec_confuseMatrix.png)

![BANTER Model Hawaii ICI Confusion
Matrix](../manuscript/manuscript_files/hawaii_ici_confuseMatrix.png)

Proximity Plot

```{r, proximity plot, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/hawaii_ec_proximity.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotProximity(banter_model_ec_RF)
dev.off()
ec_hawaii_proximityPlot <- plotProximity(banter_model_ec_RF)

png(('../manuscript/manuscript_files/hawaii_ici_proximity.png'), width = 20, height = 20, units = 'cm',  res = 300)
ici_hawaii_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_hawaii_proximityPlot <- plotProximity(banter_model_ici_RF)
```

Importance Heatmap

```{r, importance heat map, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/hawaii_ec_importance.png'), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_ec_RF, plot.type="heatmap")
dev.off()
ec_hawaii_importance <- plotImportance(banter_model_ec_RF, plot.type="heatmap")

png(('../manuscript/manuscript_files/hawaii_ici_importance.png'), width = 30, height = 25, units = 'cm',  res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_hawaii_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")
```

PlotVotes

```{r, plot votes, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/hawaii_ec_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ec_RF)
dev.off()
hawaii_votes <- plotVotes(banter_model_ec_RF)

png(('../manuscript/manuscript_files/hawaii_ici_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_hawaii_votes <- plotVotes(banter_model_ici_RF)
```

Plot Predicted Probabilities

```{r, predicted probabilities, include=TRUE, results=FALSE}
plotPredictedProbs(banter_model_ec_RF, bins=30, plot=TRUE)
plotPredictedProbs(banter_model_ici_RF, bins=30, plot=TRUE)
```

Create Figure for Publication
```{r, Publication Figure, include=TRUE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

hawaii_ec_Figure <-image_append(c(prox, heat, vote))
hawaii_ec_Figure<- image_append(c(confuse, hawaii_ec_Figure), stack=TRUE)
image_write(hawaii_ec_Figure, path = here('manuscript', 'manuscript_files','hawaii_ec_Figure.png'), format ='png')
print(hawaii_ec_Figure, info=FALSE)

confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

hawaii_ici_Figure <-image_append(c(prox, heat, vote))
hawaii_ici_Figure<- image_append(c(confuse, hawaii_ici_Figure), stack=TRUE)
image_write(hawaii_ici_Figure, path = here('manuscript', 'manuscript_files','hawaii_ici_Figure.png'), format ='png')
print(hawaii_ici_Figure, info=FALSE)
```