### 2023-06-08 manuscript/Supplement.qmd

---
title: Supplement for Open-Source machine learning BANTER acoustic classification of a cryptic echolocating species
author: Shannon Rankin
date: last-modified
output: pdf
---

```{r, load packages, include=FALSE}
library("easypackages")
libraries("tidyverse", "knitr", "kableExtra", "magick", "magrittr", "here")
here()
```

# NAtlantic
## EC Model

Detector Model:  ntree=1e4, sampsize=3; 
Event Model: ntree=1e4, sampsize=3
```{r,  NAtlantic EC Model Figure, echo=FALSE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

natlantic_ec_Figure <-image_append(c(prox, heat, vote))
natlantic_ec_Figure<- image_append(c(confuse, natlantic_ec_Figure), stack=TRUE)
print(natlantic_ec_Figure, info=FALSE)
```

## EC_ICI_ALT Model
Detector Model:  ntree=1e4, sampsize=5; 
Event Model: ntree=1e4, sampsize=5
```{r, NAtlantic EC_ICI_ALT Model Figure,  echo=FALSE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_alt_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_alt_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_alt_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_alt_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

natlantic_alt_Figure <-image_append(c(prox, heat, vote))
natlantic_alt_Figure<- image_append(c(confuse, natlantic_alt_Figure), stack=TRUE)
print(natlantic_alt_Figure, info=FALSE)
```

# SAtlantic
## EC Model
Detector Model:  ntree=1e4, sampsize=2; 
Event Model: ntree=1e5, sampsize=1
```{r, SAtlantic EC Model Figure,  echo=FALSE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

satlantic_ec_Figure <-image_append(c(prox, heat, vote))
satlantic_ec_Figure<- image_append(c(confuse, satlantic_ec_Figure), stack=TRUE)
print(satlantic_ec_Figure, info=FALSE)
```

## EC_ICI_ALT Model
Detector Model:  ntree=1e4, sampsize=4; 
Event Model: ntree=1e5, sampsize=2
```{r, SAtlantic EC_ICI_ALT Model Figure,  echo=FALSE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

satlantic_alt_Figure <-image_append(c(prox, heat, vote))
satlantic_alt_Figure<- image_append(c(confuse, satlantic_alt_Figure), stack=TRUE)
print(satlantic_alt_Figure, info=FALSE)
```

# Hawaii
## EC_ICI Model
Detector Model:  ntree=5e3, sampsize=5; 
Event Model: ntree=1e4, sampsize=4
```{r, Hawaii EC_ICI Model Figure,  echo=FALSE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

hawaii_ici_Figure <-image_append(c(prox, heat, vote))
hawaii_ici_Figure<- image_append(c(confuse, hawaii_ici_Figure), stack=TRUE)
print(hawaii_ici_Figure, info=FALSE)
```

# EPacific
## EC Model
Detector Model:  ntree=1e4, sampsize=3; 
Event Model: ntree=1e4, sampsize=4

```{r, EPacific EC Model Figure,  echo=FALSE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

epacific_ec_Figure <-image_append(c(prox, heat, vote))
epacific_ec_Figure<- image_append(c(confuse, epacific_ec_Figure), stack=TRUE)
print(epacific_ec_Figure, info=FALSE)
```