### 2023-06-08 satlantic/beakerBanter_satlantic.Rmd

---
title: 'BANTER for Beaked Whales: South Atlantic'
author: "Shannon Rankin"
date: "2023-05-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When this rmarkdown file is knit:
# FALSE means it will used saved .rds files instead of re-processing
# TRUE will run everything from scratch - this will take a long time
freshRun <- FALSE
```
###Organizational Instructions
Save this file in an R Project that includes the following foldering:
/Project
  |-Database/
  |-Binaries/
  |-Figures/
  |-satlantic
    |_ beakerBanter_satlantic.Rmd
    |_ GU1605_PG_OfflineEvents_20210809.xlsx
    |_ satlantic_banter_data.rds
    |_ satlantic_banter_data_ici.rds
    |_ satlantic_banter_model_ec_t1e4s2_t1e5s1.rds
    |_ satlantic_banter_model_ici_t1e4s4_t1e5s1.rds
    |_ satlantic_banter_model_alt_t1e4s4_t1e5s2.rds

## PAMpal Data Processing

Start by loading the required packages

```{r, load packages, message=FALSE}
library("easypackages")
libraries("PAMpal", "banter", "rfPermute", "kableExtra", "readxl", "stringr", "magick", "magrittr", "here")
here()
```

1.  **Set up our PPS** (PAMPal Settings Object) for the satlantic
    dataset

```{r, pampal settings, eval=freshRun}
pps <- PAMpalSettings(db='Database/', 
                      binaries = 'Binaries/',
                      sr_hz='auto', 
                      winLen_sec=.0025, 
                      filterfrom_khz=10, 
                      filterto_khz=NULL)
```

2.  **Process data** and save to file to eliminate repeated processing.

[If]{.underline} this is the [initial processing]{.underline}, ensure
you have set 'freshRun = TRUE' at top of this document to process and
save data. This will take some time to run.

```{r, process detections, eval=freshRun}
data <- processPgDetections(pps, mode='db', id='satlantic_bw')
saveRDS(data, 'satlantic_study.rds')
# Double check warning messages
print(getWarnings(data)$message)
```

[If]{.underline} you have already run the processing code, ensure you
have set 'freshRun = FALSE' at top of this document to read in the
existing .rds file for downstream processing.

```{r, read acoustic study, include=FALSE, eval=!freshRun}
data <- readRDS(here('satlantic', 'satlantic_study.rds'))
# Double check warning messages
print(getWarnings(data)$message)
```

3.  **Assign species identity** according to original PAMguard labels,
    then relabel for consistency across projects.

    *Final species ID for beaked whale detections was provided by A.
    DeAngelis as an excel spreadsheet. The following code chunk pulls
    out the required information (event, species) in the format required
    for PAMpal.*

```{r, assign species ID, eval=freshRun}
GU1605 <- read_excel("GU1605_PG_OfflineEvents_20210809.xlsx", sheet = "BWonly")#read in data
GU1605 <- filter(GU1605, !eventType == "BRAN")
#create formatted event/species dataframe 
speciesID <- GU1605[c("Id", "Species", "Database")] %>% 
  mutate("event" = paste(Database, Id, sep=".OE")) %>% 
  rename("species" = "Species") %>% 
  mutate("species" = str_replace_all(species, "'", ""),
         "species" = str_replace(species, "Sowerby", "Sowerbys"),
         "species" = str_replace(species, "Blainville", "Blainvilles"),
         "species" = str_replace(species, "Cuvier", "Cuviers")) %>% 
  select(all_of(c("event", "species")))

#Assign species ID to data
data <- setSpecies(data, method='manual', value = speciesID)
#Note: 238 detections are not given species IDs and will be filtered out at a later step.
```

4.  **Filter out unwanted data**: Subset and rename species for BANTER
    model

```{r, filter unwanted data, eval=freshRun}
goodSpecies<- c(c("Blainvilles", "Trues", "Cuviers", "Sowerbys", "Gervais"))
data <- filter(data, species %in% goodSpecies)
```

5.  **Calculate Inter-Click Interval** (ICI).

```{r, calculate ICI, eval=freshRun}
data <- calculateICI(data, time='peakTime')
```

6.  **Export data for BANTER** (and drop species codes that will not be
    used for training). We will create two datasets: one with ICI and
    one without ICI, and save these for import into BANTER.

```{r, banter export, eval=freshRun}
banter_data <- export_banter(data, dropSpecies = 'unid', 
                         dropVars = c('All_ici'), training=TRUE)
saveRDS(banter_data, file='satlantic_banter_data.rds')

banter_data_ici <- export_banter(data, dropSpecies = 'unid', training=TRUE)
saveRDS(banter_data_ici, file='satlantic_banter_data_ici.rds')

#save update of Acoustic Study
saveRDS(data, 'satlantic_study.rds')
```

## Build a BANTER Classification Model

### *EC (only) Model*

Initialize, Run & Evaluate Detector Model (stage 1).
```{r, banter ec detector model, eval=freshRun}
banter_model_ec <- initBanterModel(banter_data$events)
banter_model_ec <- addBanterDetector(banter_model_ec, banter_data$detectors, ntree=1e4, sampsize=2, importance = TRUE)

plotDetectorTrace(banter_model_ec, detector = paste0('Click_Detector_', 3:4))
plotDetectorTrace(banter_model_ec, detector = paste0('Click_Detector_', 5:6))
summary(banter_model_ec)
```

Run BANTER Event Model (stage 2)
```{r, banter ec event model, eval=freshRun}
banter_model_ec <- runBanterModel(banter_model_ec, ntree=1e5, sampsize=1)
summary(banter_model_ec)
```

Once a stable model is identified, save model with tree/sampsize info in the filename.
```{r,save banter ec model, eval=freshRun}
saveRDS(banter_model_ec, 'satlantic_banter_model_ec_t1e4s2_t1e5s1.rds')
```

### *ICI Model*

Initialize, Run & Evaluate Detector Model (stage 1)
```{r, banter ici detector model, eval=freshRun}
banter_model_ici <- initBanterModel(banter_data_ici$events)
banter_model_ici <- addBanterDetector(banter_model_ici, banter_data_ici$detectors, ntree=1e4, sampsize=4, importance = TRUE)

plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 3:4))
plotDetectorTrace(banter_model_ici, detector = paste0('Click_Detector_', 5:6))
summary(banter_model_ici)
```

Run BANTER Event Model (stage 2)
```{r, banter ici event model, eval=freshRun}
banter_model_ici <- runBanterModel(banter_model_ici, ntree=1e5, sampsize=1)
summary(banter_model_ici)
```

Once a stable model is identified, save model with tree/sampsize info in the filename.
```{r, save banter ici model, eval=freshRun, eval=freshRun}
saveRDS(banter_model_ici, 'satlantic_banter_model_ici_t1e4s4_t1e5s1.rds')
```

### Alternative Model (EC_ICI_ALT) with larger sampsize

Initialize, Run & Evaluate Detector Model (stage 1)
```{r, banter alt detector model, eval=freshRun}
banter_model_alt <- initBanterModel(banter_data_ici$events)
banter_model_alt <- addBanterDetector(banter_model_alt, banter_data_ici$detectors, ntree=1e4, sampsize=4, importance = TRUE)

plotDetectorTrace(banter_model_alt, detector = paste0('Click_Detector_', 3:4))
plotDetectorTrace(banter_model_alt, detector = paste0('Click_Detector_', 5:6))
summary(banter_model_alt)
```

Run BANTER Event Model (stage 2)
```{r, banter alt event model, eval=freshRun}
banter_model_alt <- runBanterModel(banter_model_alt, ntree=1e5, sampsize=2)
summary(banter_model_alt)
```

Once a stable model is identified, save model with tree/sampsize info in the filename.
```{r, save banter alt model, eval=freshRun, eval=freshRun}
saveRDS(banter_model_alt, 'satlantic_banter_model_alt_t1e4s4_t1e5s2.rds')
```

## BANTER Analytics

There are a number of visualizations/data products that allow us to
visualize our BANTER classifier; most use the rfPermute package (see
[BANTER
Guidelines](https://taikisan21.github.io/PAMpal/banterGuide.html) for
more information)

First, load the models (if not a fresh run)
```{r, read banter models, include=FALSE, eval=!freshRun}
banter_model_ec <- readRDS(here('satlantic', 'satlantic_banter_model_ec_t1e4s2_t1e5s1.rds'))
banter_model_ici <- readRDS(here('satlantic', 'satlantic_banter_model_ici_t1e4s4_t1e5s1.rds'))
banter_model_alt <- readRDS(here('satlantic', 'satlantic_banter_model_alt_t1e4s4_t1e5s2.rds'))
```

Identify the model you would like to examine (comment out the
model you [do not]{.underline} want to examine).

```{r, identify model}
model_ec <- banter_model_ec
modelname_ec <- "banter_model_ec"

model_ici <- banter_model_ici
modelname_ici <- "banter_model_ici"

model_alt <- banter_model_alt
modelname_alt <- "banter_model_alt"
```

Extract the Random Forest model object from our BANTER model for
analysis.

```{r, extract RF model, include=TRUE}
banter_model_ec_RF <- getBanterModel(model_ec)
banter_model_ici_RF <- getBanterModel(model_ici)
banter_model_alt_RF <- getBanterModel(model_alt)
```

Class Priors (Expected Error Rate)

```{r, class priors, include=TRUE}
satlantic_ec_priors <- classPriors(banter_model_ec_RF, NULL)[,1]
satlantic_ici_priors <- classPriors(banter_model_ici_RF, NULL)[,1]
satlantic_alt_priors <- classPriors(banter_model_alt_RF, NULL)[,1]
```

Confusion Matrix (simple model)

```{r, confusion matrix, include=TRUE}
satlantic_ec_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ec_RF)
satlantic_ec_confuseMatrix <- cbind(satlantic_ec_confuseMatrix, priors = satlantic_ec_priors)
satlantic_ec_confuseMatrix <- kable(satlantic_ec_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(6, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(6, bold = TRUE)%>%
  save_kable('../manuscript/manuscript_files/satlantic_ec_confuseMatrix.png', zoom = 9)

satlantic_ici_confuseMatrix <- rfPermute::confusionMatrix(banter_model_ici_RF)
satlantic_ici_confuseMatrix <- cbind(satlantic_ici_confuseMatrix, priors = satlantic_ici_priors)
satlantic_ici_confuseMatrix <- kable(satlantic_ici_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(6, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(6, bold = TRUE)%>%
  save_kable('../manuscript/manuscript_files/satlantic_ici_confuseMatrix.png', zoom = 9)

satlantic_alt_confuseMatrix <- rfPermute::confusionMatrix(banter_model_alt_RF)
satlantic_alt_confuseMatrix <- cbind(satlantic_alt_confuseMatrix, priors = satlantic_alt_priors)
satlantic_alt_confuseMatrix <- kable(satlantic_alt_confuseMatrix, align = "c", digits = c(0,0,0,0,2,2,2))%>%
  kable_classic()%>%
  column_spec(5, border_right = TRUE)%>%
  row_spec(0, bold = TRUE)%>%
  row_spec(3,hline_after = TRUE)%>%
  row_spec(5, bold = TRUE)%>%
  save_kable('../manuscript/manuscript_files/satlantic_alt_confuseMatrix.png', zoom = 9)

```

![BANTER Model S.Atlantic EC Confusion
Matrix](../manuscript/manuscript_files/satlantic_ec_confuseMatrix.png)

![BANTER Model S.Atlantic ICI Confusion
Matrix](../manuscript/manuscript_files/satlantic_ici_confuseMatrix.png)

![BANTER Model S.Atlantic Alternative Model Confusion
Matrix](../manuscript/manuscript_files/satlantic_alt_confuseMatrix.png)

Proximity Plot

```{r, proximity plot, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/satlantic_ec_proximity.png'), width = 20, height = 20, units = 'cm', res = 300)
satlantic_ec_proximityPlot <- plotProximity(banter_model_ec_RF)
dev.off()
satlantic_ec_proximityPlot <- plotProximity(banter_model_ec_RF)

png(('../manuscript/manuscript_files/satlantic_ici_proximity.png'), width = 20, height = 20, units = 'cm', res = 300)
ici_satlantic_proximityPlot <- plotProximity(banter_model_ici_RF)
dev.off()
ici_satlantic_proximityPlot <- plotProximity(banter_model_ici_RF)

png(('../manuscript/manuscript_files/satlantic_alt_proximity.png'), width = 20, height = 20, units = 'cm', res = 300)
alt_satlantic_proximityPlot <- plotProximity(banter_model_alt_RF)
dev.off()
alt_satlantic_proximityPlot <- plotProximity(banter_model_alt_RF)
```

Importance Heatmap

```{r, importance heat map, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/satlantic_ec_importance.png'), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_ec_RF, plot.type="heatmap")
dev.off()
satlantic_importance <- plotImportance(banter_model_ec_RF, plot.type="heatmap")

png(('../manuscript/manuscript_files/satlantic_ici_importance.png'), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_ici_RF, plot.type="heatmap")
dev.off()
ici_satlantic_importance <- plotImportance(banter_model_ici_RF, plot.type="heatmap")

png(('../manuscript/manuscript_files/satlantic_alt_importance.png'), width = 30, height = 25, units = 'cm', res = 300)
plotImportance(banter_model_alt_RF, plot.type="heatmap")
dev.off()
alt_satlantic_importance <- plotImportance(banter_model_alt_RF, plot.type="heatmap")
```

PlotVotes

```{r, plot votes, include=TRUE, results=FALSE}
png(('../manuscript/manuscript_files/satlantic_ec_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ec_RF)
dev.off()
satlantic_votes <- plotVotes(banter_model_ec_RF)

png(('../manuscript/manuscript_files/satlantic_ici_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_ici_RF)
dev.off()
ici_satlantic_votes <- plotVotes(banter_model_ici_RF)

png(('../manuscript/manuscript_files/satlantic_alt_votes.png'), width = 20, height = 20, units = 'cm',  res = 300)
plotVotes(banter_model_alt_RF)
dev.off()
alt_satlantic_votes <- plotVotes(banter_model_alt_RF)
```

Plot Predicted Probabilities

```{r, predicted probabilities, include=TRUE, results=FALSE}
plotPredictedProbs(banter_model_ec_RF, bins=30, plot=TRUE)
plotPredictedProbs(banter_model_ici_RF, bins=30, plot=TRUE)
plotPredictedProbs(banter_model_alt_RF, bins=30, plot=TRUE)
```

Create Figure for Publication
```{r, Publication Figure, include=TRUE}
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

satlantic_ec_Figure <-image_append(c(prox, heat, vote))
satlantic_ec_Figure<- image_append(c(confuse, satlantic_ec_Figure), stack=TRUE)
image_write(satlantic_ec_Figure, path = here('manuscript', 'manuscript_files','satlantic_ec_Figure.png'), format ='png')
print(satlantic_ec_Figure, info=FALSE)

confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ici_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ici_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ici_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ici_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

satlantic_ici_Figure <-image_append(c(prox, heat, vote))
satlantic_ici_Figure<- image_append(c(confuse, satlantic_ici_Figure), stack=TRUE)
image_write(satlantic_ici_Figure, path = here('manuscript', 'manuscript_files','satlantic_ici_Figure.png'), format ='png')
print(satlantic_ici_Figure, info=FALSE)

confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "50x130")%>%
  image_annotate("a) Confusion Matrix", size=100, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_votes.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("d) Vote Plot", size=100, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_annotate("b) Proximity Plot", size=100, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_importance.png'))%>%
  image_border(color="#ffffff", geometry = "270x130")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=100, color = "black")

satlantic_alt_Figure <-image_append(c(prox, heat, vote))
satlantic_alt_Figure<- image_append(c(confuse, satlantic_alt_Figure), stack=TRUE)
image_write(satlantic_alt_Figure, path = here('manuscript', 'manuscript_files','satlantic_alt_Figure.png'), format ='png')
print(satlantic_alt_Figure, info=FALSE)
```