### 2024-02-02 manuscript/Supplement.qmd
Final Revision for Manuscript
Revise mansucript including supplement, figures, etc. This is as submitted.

---
title: Supplement for Open-Source machine learning BANTER acoustic classification of beaked whale echolocation pulses
author: Shannon Rankin
date: last-modified
#output: pdf
format: docx
---

```{r, load packages, include=FALSE}
library("easypackages")
libraries("tidyverse", "knitr", "kableExtra", "magick", "magrittr", "ggplot2", "here")
here()
```

# NAtlantic
Results for the NAtlantic EC (only) model are provided in @fig-natlantic_EC (Detector Model: *sampsize* = 3, *ntree* = 10,000; Event Model: *sampsize* = 3, *ntree* = 10,000).

```{r,  NAtlantic EC Model Figure, echo=FALSE}
#| label: fig-natlantic_EC
#| fig-cap: BANTER classification results from the NAtlantic dataset including echolocation pulses (EC). Confusion matrix (a) provides the percent correct classification for each species (pct.correct), lower confidence intervals (LCI_0.95), upper confidence intervals (UCI_0.95), and priors (expected error rate). Proximity plot (b) for species events from BANTER model (central dot color represents true species identity; color of circle surrounding dot represents BANTER species classification). Heat map (c) for ranks of ten most important variables; colors scale from most important predictors (dark red) to least important predictors (dark blue). Vote Plot (d) shows the vote distribution for each event (vertical slice) for each species; distribution of votes by species is shown by their representative color.
#| #| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| out-width: 100%
#| echo: false
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "80x208")%>%
  image_annotate("a) Confusion Matrix", size=120, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("d) Vote Plot", size=120, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("b) Proximity Plot", size=120, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'natlantic_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "110x260")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=120, color = "black")

natlantic_ec_Figure <-image_append(c(prox, heat, vote))
natlantic_ec_Figure<- image_append(c(confuse, natlantic_ec_Figure), stack=TRUE)
print(natlantic_ec_Figure, info=FALSE)
```

# SAtlantic

Results for the SAtlantic EC (only) model are provided in @fig-satlantic_EC (Detector Model: *sampsize* = 8, *ntree* = 10,000; Event Model: *sampsize* = 1, *ntree* = 100,000).

```{r, SAtlantic EC Model Figure,  echo=FALSE}
#| label: fig-satlantic_EC
#| fig-cap: BANTER classification results from the SAtlantic dataset including echolocation pulses (EC). Confusion matrix (a) provides the percent correct classification for each species (pct.correct), lower confidence intervals (LCI_0.95), upper confidence intervals (UCI_0.95), and priors (expected error rate). Proximity plot (b) for species events from BANTER model (central dot color represents true species identity; color of circle surrounding dot represents BANTER species classification). Heat map (c) for ranks of ten most important variables; colors scale from most important predictors (dark red) to least important predictors (dark blue). Vote Plot (d) shows the vote distribution for each event (vertical slice) for each species; distribution of votes by species is shown by their representative color.
#| #| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| out-width: 100%
#| echo: false
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "80x208")%>%
  image_annotate("a) Confusion Matrix", size=120, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("d) Vote Plot", size=120, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("b) Proximity Plot", size=120, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "110x260")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=120, color = "black")

satlantic_ec_Figure <-image_append(c(prox, heat, vote))
satlantic_ec_Figure<- image_append(c(confuse, satlantic_ec_Figure), stack=TRUE)
print(satlantic_ec_Figure, info=FALSE)
```

Results for the SAtlantic EC_IPI_ALT model are provided in @fig-satlantic_EC_IPI_ALT (Detector Model: *sampsize* = 4, *ntree* = 10,000; Event Model: *sampsize* = 3, *ntree* = 100,000).

```{r, SAtlantic EC_IPI_ALT Model Figure,  echo=FALSE}
#| label: fig-satlantic_EC_IPI_ALT
#| fig-cap: BANTER classification results from the alternative SAtlantic dataset including echolocation pulses and inter-pulse interval with a large sampsize (EC_IPI_ALT). Confusion matrix (a) provides the percent correct classification for each species (pct.correct), lower confidence intervals (LCI_0.95), upper confidence intervals (UCI_0.95), and priors (expected error rate). Proximity plot (b) for species events from BANTER model (central dot color represents true species identity; color of circle surrounding dot represents BANTER species classification). Heat map (c) for ranks of ten most important variables; colors scale from most important predictors (dark red) to least important predictors (dark blue). Vote Plot (d) shows the vote distribution for each event (vertical slice) for each species; distribution of votes by species is shown by their representative color.
#| #| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| out-width: 100%
#| echo: false
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "80x208")%>%
  image_annotate("a) Confusion Matrix", size=120, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_votes.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("d) Vote Plot", size=120, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("b) Proximity Plot", size=120, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'satlantic_alt_importance.png'))%>%
  image_border(color="#ffffff", geometry = "110x260")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=120, color = "black")

satlantic_alt_Figure <-image_append(c(prox, heat, vote))
satlantic_alt_Figure<- image_append(c(confuse, satlantic_alt_Figure), stack=TRUE)
print(satlantic_alt_Figure, info=FALSE)
```

# Hawaii

Results for the Hawaii EC_IPI model are provided in @fig-hawaii_EC_IPI (Detector Model: *sampsize* = 5, *ntree* = 100,000; Event Model: *sampsize* = 4, *ntree* = 10,000).

```{r, Hawaii EC_IPI Model Figure,  echo=FALSE}
#| label: fig-hawaii_EC_IPI
#| fig-cap: BANTER classification results from the Hawaii dataset including echolocation pulses and inter-pulse interval (EC_IPI). Confusion matrix (a) provides the percent correct classification for each species (pct.correct), lower confidence intervals (LCI_0.95), upper confidence intervals (UCI_0.95), and priors (expected error rate). Proximity plot (b) for species events from BANTER model (central dot color represents true species identity; color of circle surrounding dot represents BANTER species classification). Heat map (c) for ranks of ten most important variables; colors scale from most important predictors (dark red) to least important predictors (dark blue). Vote Plot (d) shows the vote distribution for each event (vertical slice) for each species; distribution of votes by species is shown by their representative color.
#| #| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| out-width: 100%
#| echo: false
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "80x208")%>%
  image_annotate("a) Confusion Matrix", size=120, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_votes.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("d) Vote Plot", size=120, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("b) Proximity Plot", size=120, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'hawaii_ici_importance.png'))%>%
  image_border(color="#ffffff", geometry = "110x260")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=120, color = "black")

hawaii_ici_Figure <-image_append(c(prox, heat, vote))
hawaii_ici_Figure<- image_append(c(confuse, hawaii_ici_Figure), stack=TRUE)
print(hawaii_ici_Figure, info=FALSE)
```

# EPacific

Results for the EPacific EC (only) model are provided in @fig-epacific_EC (Detector Model: *sampsize* = 4, *ntree* = 10,000; Event Model: *sampsize* = 4, *ntree* = 10,000).

```{r, EPacific EC Model Figure,  echo=FALSE}
#| label: fig-epacific_EC
#| fig-cap: BANTER classification results from the EPacific dataset including echolocation pulses (EC). Confusion matrix (a) provides the percent correct classification for each species (pct.correct), lower confidence intervals (LCI_0.95), upper confidence intervals (UCI_0.95), and priors (expected error rate). Proximity plot (b) for species events from BANTER model (central dot color represents true species identity; color of circle surrounding dot represents BANTER species classification). Heat map (c) for ranks of ten most important variables; colors scale from most important predictors (dark red) to least important predictors (dark blue). Vote Plot (d) shows the vote distribution for each event (vertical slice) for each species; distribution of votes by species is shown by their representative color.
#| #| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| out-width: 100%
#| echo: false
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "80x208")%>%
  image_annotate("a) Confusion Matrix", size=120, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_votes.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("d) Vote Plot", size=120, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("b) Proximity Plot", size=120, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ec_importance.png'))%>%
  image_border(color="#ffffff", geometry = "110x260")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=120, color = "black")

epacific_ec_Figure <-image_append(c(prox, heat, vote))
epacific_ec_Figure<- image_append(c(confuse, epacific_ec_Figure), stack=TRUE)
print(epacific_ec_Figure, info=FALSE)
```

Results for the EPacific EC (only) model are provided in @fig-epacific_EC_IPI (Detector Model: *sampsize* = 3, *ntree* = 10,000; Event Model: *sampsize* = 4, *ntree* = 10,000).

```{r, EPacific EC IPI Model Figure,  echo=FALSE}
#| label: fig-epacific_EC_IPI
#| fig-cap: BANTER classification results from the EPacific dataset including echolocation pulses and inter-pulse interval (EC_IPI). Confusion matrix (a) provides the percent correct classification for each species (pct.correct), lower confidence intervals (LCI_0.95), upper confidence intervals (UCI_0.95), and priors (expected error rate). Proximity plot (b) for species events from BANTER model (central dot color represents true species identity; color of circle surrounding dot represents BANTER species classification). Heat map (c) for ranks of ten most important variables; colors scale from most important predictors (dark red) to least important predictors (dark blue). Vote Plot (d) shows the vote distribution for each event (vertical slice) for each species; distribution of votes by species is shown by their representative color.
#| #| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| out-width: 100%
#| echo: false
confuse <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_confuseMatrix.png'))%>%
  image_border(color="#ffffff", geometry = "80x208")%>%
  image_annotate("a) Confusion Matrix", size=120, color = "black")
vote <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_votes.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("d) Vote Plot", size=120, color = "black")
prox <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_proximity.png'))%>%
  image_border(color="#ffffff", geometry = "260x125")%>%
  image_annotate("b) Proximity Plot", size=120, color = "black")
heat <- magick::image_read(here('manuscript', 'manuscript_files', 'epacific_ici_importance.png'))%>%
  image_border(color="#ffffff", geometry = "110x260")%>%
  image_scale("3300")%>%
  image_annotate("d) Importance Heat Map", size=120, color = "black")

epacific_ici_Figure <-image_append(c(prox, heat, vote))
epacific_ici_Figure<- image_append(c(confuse, epacific_ici_Figure), stack=TRUE)
print(epacific_ici_Figure, info=FALSE)
```